---
title: "Main Research"
output: html_notebook
---

By Jack Wilburn

### Libraries

```{r, message = FALSE, warning = FALSE}
library(readr)
library(mosaic)
library(gridExtra)
library(caret)
library(reshape)
library(car)
library(broom)
library(e1071)
library(leaps)
library(psych)
```

### Importing and Cleaning Data


```{r, message = FALSE, warning = FALSE}
# execs import and cleaning
execs = read_csv("SeniorExecs.csv")
execs = execs[,c(1,2,4:37,44:54)]
execs$dimIPart1 = execs$DIMI1
execs$dimEPart1 = execs$DIME1
execs$dimSPart1 = execs$DIMS1
execs$intIPart1 = execs$INTI1
execs$intEPart1 = execs$INTE1
execs$intSPart1 = execs$INTS1
execs$difPart1 = execs$DIF1
execs$intPart1 = execs$INT1
execs$dimPart1 = execs$DIM1
execs$diPart1 = execs$DI1
execs$disPart1 = execs$DIS1
execs$VQ1 = execs$VQ1
execs$VQ2 = execs$VQ2
execs$dimPercPart1 = execs$DIMper1
execs$intPercPart1 = execs$INTper1
execs$aiPercPart1 = execs$AIper1
execs$dimIPart2 = execs$DIMI2
execs$dimEPart2 = execs$DIME2
execs$dimSPart2 = execs$DIMS2
execs$intIPart2 = execs$INTI2
execs$intEPart2 = execs$INTE2
execs$intSPart2 = execs$INTS2
execs$difPart2 = execs$DIF2
execs$intPart2 = execs$INT2
execs$dimPart2 = execs$DIM2
execs$diPart2 = execs$DI2
execs$disPart2 = execs$DIS2
execs$SQ1 = execs$SQ1
execs$SQ2 = execs$SQ2
execs$dimPercPart2 = execs$DIMper2
execs$intPercPart2 = execs$INTper2
execs$aiPercPart2 = execs$AIper2
execs$BQr1 = execs$BQr1
execs$BQr2 = execs$BQr2
execs$BQa1 = execs$BQa1
execs$BQa2 = execs$BQa2
execs$CQ1 = execs$CQ1
execs$CQ2 = execs$CQ2
execs$rho1 = (execs$Rho1)/1000
execs$rho2 = (execs$Rho2)/1000
execs$difPart1difPart2 = execs$DIF1divDIF2
execs$type = "Executive"
execs = cbind(execs[,48:58], execs[,38:39], execs[59:72], execs[,40:41], execs[,73:75], execs[,42:47], execs[,76:79])
execs[, 1:41] = sapply(execs[, 1:41], as.numeric)
execs$type = factor(execs$type)

# colStus import and cleaning
colStus = read_csv("CollegeStudents.csv")
colStus = colStus[,c(14:15,17:50,57:67)]
colStus$dimIPart1 = colStus$DIMI1
colStus$dimEPart1 = colStus$DIME1
colStus$dimSPart1 = colStus$DIMS1
colStus$intIPart1 = colStus$INTI1
colStus$intEPart1 = colStus$INTE1
colStus$intSPart1 = colStus$INTS1
colStus$difPart1 = colStus$DIF1
colStus$intPart1 = colStus$INT1
colStus$dimPart1 = colStus$DIM1
colStus$diPart1 = colStus$DI1
colStus$disPart1 = colStus$DIS1
colStus$VQ1 = colStus$VQ1
colStus$VQ2 = colStus$VQ2
colStus$dimPercPart1 = colStus$DIMper1
colStus$intPercPart1 = colStus$INTper1
colStus$aiPercPart1 = (colStus$AIper1)
colStus$dimIPart2 = colStus$DIMI2
colStus$dimEPart2 = colStus$DIME2
colStus$dimSPart2 = colStus$DIMS2
colStus$intIPart2 = colStus$INTI2
colStus$intEPart2 = colStus$INTE2
colStus$intSPart2 = colStus$INTS2
colStus$difPart2 = colStus$DIF2
colStus$intPart2 = colStus$INT2
colStus$dimPart2 = colStus$DIM2
colStus$diPart2 = colStus$DI2
colStus$disPart2 = colStus$DIS2
colStus$SQ1 = colStus$SQ1
colStus$SQ2 = colStus$SQ2
colStus$dimPercPart2 = colStus$DIMper2
colStus$intPercPart2 = colStus$INTper2
colStus$aiPercPart2 = (colStus$AIper2)
colStus$BQr1 = colStus$BQr1
colStus$BQr2 = colStus$BQR2
colStus$BQa1 = colStus$BQa1
colStus$BQa2 = colStus$BQa2
colStus$CQ1 = colStus$CQ1
colStus$CQ2 = colStus$CQ2
colStus$rho1 = colStus$Rho1
colStus$rho2 = colStus$Rho2
colStus$difPart1difPart2 = colStus$DIF1divDIF2
colStus$type = "CollegeStudent"
colStus = cbind(colStus[,48:58], colStus[,38:39], colStus[59:72], colStus[,40:41], colStus[,73:75], colStus[,42], colStus[,76], colStus[,44:47], colStus[,77:80])
colStus[, 1:41] = sapply(colStus[, 1:41], as.numeric)
colStus$type = factor(colStus$type)
colStus = na.omit(colStus)
colStus$aiPercPart1 = (colStus$aiPercPart1) /100
colStus$aiPercPart2 = (colStus$aiPercPart2) /100
colStus$dimPercPart1 = (colStus$dimPercPart1) /100
colStus$dimPercPart2 = (colStus$dimPercPart2) /100
colStus$intPercPart1 = (colStus$intPercPart1) /100
colStus$intPercPart2 = (colStus$intPercPart2) /100
colStus$difPart1difPart2 = (colStus$difPart1difPart2) / 100

# entreps import and cleaning
entreps = read_csv("Entrepreneurs.csv")
entreps = entreps[,c(15:51, 58:68)]
entreps$dimIPart1 = entreps$DIMI1
entreps$dimEPart1 = entreps$DIME1
entreps$dimSPart1 = entreps$DIMS1
entreps$intIPart1 = entreps$INTI1
entreps$intEPart1 = entreps$INTE1
entreps$intSPart1 = entreps$INTS1
entreps$difPart1 = entreps$DIF1
entreps$intPart1 = entreps$INT1
entreps$dimPart1 = entreps$DIM1
entreps$diPart1 = entreps$DI1
entreps$disPart1 = entreps$DIS1
entreps$VQ1 = entreps$VQ1
entreps$VQ2 = entreps$VQ2
entreps$dimPercPart1 = entreps$DIMper1
entreps$intPercPart1 = entreps$INTper1
entreps$aiPercPart1 = (entreps$AIper1)
entreps$dimIPart2 = entreps$DIMI2
entreps$dimEPart2 = entreps$DIME2
entreps$dimSPart2 = entreps$DIMS2
entreps$intIPart2 = entreps$INTI2
entreps$intEPart2 = entreps$INTE2
entreps$intSPart2 = entreps$INTS2
entreps$difPart2 = entreps$DIF2
entreps$intPart2 = entreps$INT2
entreps$dimPart2 = entreps$DIM2
entreps$diPart2 = entreps$DI2
entreps$disPart2 = entreps$DIS2
entreps$SQ1 = entreps$SQ1
entreps$SQ2 = entreps$SQ2
entreps$dimPercPart2 = entreps$DIMper2
entreps$intPercPart2 = entreps$INTper2
entreps$aiPercPart2 = (entreps$AIper2)
entreps$BQr1 = entreps$BQr1
entreps$BQr2 = entreps$BQR2
entreps$BQa1 = entreps$BQa1
entreps$BQa2 = entreps$BQa2
entreps$CQ1 = entreps$CQ1
entreps$CQ2 = entreps$CQ2
entreps$rho1 = entreps$Rho1
entreps$rho2 = entreps$Rho2
entreps$difPart1difPart2 = entreps$DIF1divDIF2
entreps$type = "Entrepreneur"
entreps = cbind(entreps[,49:59], entreps[,39:40], entreps[60:73], entreps[,41:42], entreps[,74:76], entreps[,43], entreps[,77], entreps[,45:48], entreps[,78:81])
entreps[, 1:41] = sapply(entreps[, 1:41], as.numeric)
entreps$type = factor(entreps$type)
entreps$aiPercPart1 = (entreps$aiPercPart1) /100
entreps$aiPercPart2 = (entreps$aiPercPart2) /100
entreps$dimPercPart1 = (entreps$dimPercPart1) /100
entreps$dimPercPart2 = (entreps$dimPercPart2) /100
entreps$intPercPart1 = (entreps$intPercPart1) /100
entreps$intPercPart2 = (entreps$intPercPart2) /100
entreps$difPart1difPart2 = (entreps$difPart1difPart2) / 100

# senMans importing and cleaning
senMans = read_csv("SeniorMgrs.csv")
senMans = senMans[,c(15:16, 18:51, 58:68)]
senMans$dimIPart1 = senMans$DIMI1
senMans$dimEPart1 = senMans$DIME1
senMans$dimSPart1 = senMans$DIMS1
senMans$intIPart1 = senMans$INTI1
senMans$intEPart1 = senMans$INTE1
senMans$intSPart1 = senMans$INTS1
senMans$difPart1 = senMans$DIF1
senMans$intPart1 = senMans$INT1
senMans$dimPart1 = senMans$DIM1
senMans$diPart1 = senMans$DI1
senMans$disPart1 = senMans$DIS1
senMans$VQ1 = senMans$VQ1
senMans$VQ2 = senMans$VQ2
senMans$dimPercPart1 = senMans$DIMper1
senMans$intPercPart1 = senMans$INTper1
senMans$aiPercPart1 = (senMans$AIper1)
senMans$dimIPart2 = senMans$DIMI2
senMans$dimEPart2 = senMans$DIME2
senMans$dimSPart2 = senMans$DIMS2
senMans$intIPart2 = senMans$INTI2
senMans$intEPart2 = senMans$INTE2
senMans$intSPart2 = senMans$INTS2
senMans$difPart2 = senMans$DIF2
senMans$intPart2 = senMans$INT2
senMans$dimPart2 = senMans$DIM2
senMans$diPart2 = senMans$DI2
senMans$disPart2 = senMans$DIS2
senMans$SQ1 = senMans$SQ1
senMans$SQ2 = senMans$SQ2
senMans$dimPercPart2 = senMans$DIMper2
senMans$intPercPart2 = senMans$INTper2
senMans$aiPercPart2 = (senMans$AIper2)
senMans$BQr1 = senMans$BQr1
senMans$BQr2 = senMans$BQr2
senMans$BQa1 = senMans$BQa1
senMans$BQa2 = senMans$BQa2
senMans$CQ1 = senMans$CQ1
senMans$CQ2 = senMans$CQ2
senMans$rho1 = senMans$Rho1
senMans$rho2 = senMans$Rho2
senMans$difPart1difPart2 = senMans$DIF1divDIF2
senMans$type = "SeniorManager"
senMans = cbind(senMans[,48:58], senMans[,38:39], senMans[59:72], senMans[,40:41], senMans[,73:75], senMans[,42:47], senMans[,76:79])
senMans[, 1:41] = sapply(senMans[, 1:41], as.numeric)
senMans$type = factor(senMans$type)
senMans = na.omit(senMans)
senMans$aiPercPart1 = (senMans$aiPercPart1) /100
senMans$aiPercPart2 = (senMans$aiPercPart2) /100
senMans$dimPercPart1 = (senMans$dimPercPart1) /100
senMans$dimPercPart2 = (senMans$dimPercPart2) /100
senMans$intPercPart1 = (senMans$intPercPart1) /100
senMans$intPercPart2 = (senMans$intPercPart2) /100
senMans$difPart1difPart2 = (senMans$difPart1difPart2) / 100

# Binding datasets together to analyze and draw out differences between groups
comb = rbind(execs, colStus, entreps, senMans)
comb[, 1:41] = sapply(comb[, 1:41], as.numeric)
comb$type = factor(comb$type)
```


### EDA

```{r, fig.height = 6}
a = ggplot() +
  geom_density(aes(x = execs$CQ1), alpha = 0.5, fill = "blue") +
  geom_density(aes(x = colStus$CQ1), alpha = 0.5, fill = "green") + 
  geom_density(aes(x = entreps$CQ1), alpha = 0.5, fill = "red") +
  geom_density(aes(x = senMans$CQ1), alpha = 0.5, fill = "black") +
  labs(x = "CQ1")
b = ggplot() +
  geom_density(aes(x = execs$rho1), alpha = 0.5, fill = "blue") +
  geom_density(aes(x = colStus$rho1), alpha = 0.5, fill = "green") + 
  geom_density(aes(x = entreps$rho1), alpha = 0.5, fill = "red") +
  geom_density(aes(x = senMans$rho1), alpha = 0.5, fill = "black") +
  labs(x = "Rho1")
c = ggplot() +
  geom_density(aes(x = execs$rho2), alpha = 0.5, fill = "blue") +
  geom_density(aes(x = colStus$rho2), alpha = 0.5, fill = "green") + 
  geom_density(aes(x = entreps$rho2), alpha = 0.5, fill = "red") +
  geom_density(aes(x = senMans$rho2), alpha = 0.5, fill = "black") +
  labs(x = "Rho2")
e = ggplot() +
  geom_density(aes(x = execs$aiPercPart1), alpha = 0.5, fill = "blue") +
  geom_density(aes(x = colStus$aiPercPart1), alpha = 0.5, fill = "green") + 
  geom_density(aes(x = entreps$aiPercPart1), alpha = 0.5, fill = "red") +
  geom_density(aes(x = senMans$aiPercPart1), alpha = 0.5, fill = "black") +
  labs(x = "AI Percentage Part 1")
f = ggplot() +
  geom_density(aes(x = execs$aiPercPart2), alpha = 0.5, fill = "blue") +
  geom_density(aes(x = colStus$aiPercPart2), alpha = 0.5, fill = "green") + 
  geom_density(aes(x = entreps$aiPercPart2), alpha = 0.5, fill = "red") +
  geom_density(aes(x = senMans$aiPercPart2), alpha = 0.5, fill = "black") +
  labs(x = "AI Percentage Part 2")
grid.arrange(a, b, c, e, f, ncol = 2)
```

### Questions

1) Are rho1 and rho2 correlated in practice?
2) do rho1 scores between samples differ
3) Take samples of execs to examine between categories.
4) Is the correlation that's baked into the model present in corelations

### Testing

1) 
```{r}
ggplot(comb, aes(rho1, rho2)) + geom_point(alpha = 0.5, aes(color = type))
cor(comb$rho1, comb$rho2)
```

2)

```{r}
sample1 = sample(nrow(execs), 94)
sampledexces1 = execs[sample1,]
sample2 = sample(nrow(execs), 94)
sampledexces2 = execs[sample2,]
sample3 = sample(nrow(execs), 94)
sampledexces3 = execs[sample3,]

t.test(sampledexces1$rho1, entreps$rho1)
t.test(sampledexces2$rho1, entreps$rho1)
t.test(sampledexces3$rho1, entreps$rho1)
```


3)

```{r}
pcaentreps = prcomp(entreps[1:41], scale = TRUE, center = TRUE)
pcaentreps$rotation[1:41,1:7]
biplot(pcaentreps)

entsenman = rbind(entreps, senMans)
pcaentsenman = prcomp(entsenman[1:41], scale = TRUE, center = TRUE)
summary(pcaentsenman)

ggplot(as.data.frame(pcaentsenman$x), aes(PC1, PC2)) + 
  geom_point(aes(color = entsenman$type))
```

Significantly different variables: dimEPart1, dipart1, VQ1, VQ2, dimpercpart1, dimIpart2, intSPart2, BQr1, BQa1, BQa2, CQ1, CQ2, rho1, rho2, difPart1difPart2

### Entreps vs SenMans

```{r, fig.width = 10, fig.height=20}
entsenman = rbind(entreps, senMans)
meltentsenman = melt(entsenman)
ggplot(meltentsenman) +
  geom_density(aes(x = value, fill = type), alpha = 0.5) +
  facet_wrap( ~ variable, scales = "free", ncol = 4)
```

Significantly different variables: dimEPart1, dispart1, VQ1, VQ2, dimpercpart1, dimIpart2, intSPart2, BQr1, BQa1, BQa2, CQ1, CQ2, rho1, rho2, difPart1difPart2

```{r}
set.seed(10)
sample = sample(nrow(entsenman), nrow(entsenman)/2)
train = entsenman[sample,]
test = entsenman[-sample,]

glm = glm(type ~ dimEPart1 + diPart1 + VQ1 + VQ2 + dimPercPart1 + dimIPart2 + intSPart2 + BQr1 + BQa1 + BQa2 + CQ1 + CQ2 + rho1 + rho2 + difPart1difPart2, data = entsenman, family = "binomial")
Anova(glm)
anova(glm)

glm2 = glm(type ~ dimIPart2 + BQr1 + CQ1 + rho1 + rho2, data = entsenman, family = "binomial")
Anova(glm2)
anova(glm2)

glm2 = glm(type ~ dimIPart2 + BQr1 + CQ1 + rho1 + rho2, data = train, family = "binomial")
preds = predict(glm2, test, response = "binomial")
preds = ifelse(preds>0, "SeniorManager", "Entrepreneur")
confusionMatrix(preds, test$type)

glm3 = glm(type ~ . - difPart1 - dimPart1 - VQ2 - intPart2 - SQ1 - SQ2 - aiPercPart1 - intEPart2 - difPart2
           - intIPart1 - intEPart1 - intSPart1 - intEPart1,
           data = entsenman, family = "binomial")
Anova(glm3)
anova(glm3)

glm3 = glm(type ~ ., data = train, family = "binomial")
preds = predict(glm3, test, response = "binomial")
preds = ifelse(preds>0, "SeniorManager", "Entrepreneur")
confusionMatrix(preds, test$type)
```

feature selection

```{r}
ffs = regsubsets(type ~ ., data = entsenman, nvmax = 10)
summary(ffs)
plot(ffs)
```

Variables where the 2 groups differ:
```{r}
# differ
print("Variables that differ between groups:")
for (i in names(entsenman)[1:41]) {
  a = wilcox.test(get(i) ~ type, data = entsenman)
  if (a$p.value < 0.05) {
    print(c(i, a$p.value))
  }
}
print(" ")

# don't differ
print("Variables that don't differ between groups:")
for (i in names(entsenman)[1:41]) {
  a = wilcox.test(get(i) ~ type, data = entsenman)
  if (a$p.value > 0.05) {
    print(c(i, a$p.value))
  }
}
```

EFA in R

```{r}
covMat = cov(comb[1:41])
covMat
### some covariance is too high, need to remove variables that covary too much to run factor analysis!
factors = factanal(covmat = covMat, factors = 7, n.obs = nrow(comb), rotation = "varimax")
factors
```





















